---
title: "R Notebook"
output: html_notebook
---

Calculating and comparing residencies for white sturgeon and chinook salmon, 2012-2016.

Detection years run from July to June (i.e. July 2011 through June 2012).  We have data through detection year 2015 (July 2015-June 2016).

There are estimates for the following groups and years:

White Sturgeon: four years of estimates for those 67 fish tagged in 2012, and two years of estimates for the 25 fish tagged in 2014.  These combine to make four years of estimates total, with varying numbers of returning fish.

Chinook salmon: four years of estimates for the fish tagged in fall 2012, 2013, 2014, and 2015.

```{r}
#  Residence Time #
library(ybp)
library(fishtrackr)
library(tidyverse)

a <- alltags
groups <- split(a, a$`Tagging Group`)
g <- map(groups, select, TagID) # selects only the TagID column of each dataframe
tags <- unlist(g, recursive = FALSE) # have to unlist the first level to be able to manipulate with these as vectors

d <- all69khz_grouped # all detections, with grouped gated receivers

# we'll do white sturgeon first
d3 <- filter(d, TagID %in% tags$wst_2012.TagID | TagID %in% tags$wst_2014.TagID) # contains all the detections for both tagging groups of white sturgeon - still have to filter further by year to get residence

d3 <- d3 %>% 
  group_by(TagID) %>% 
  filter(DateTimeUTC >= DateTagged) %>% 
  arrange(DateTimeUTC)  # make sure we don't get any detections for individual fish before they were tagged


# White sturgeon residence in detyear 2011 (which includes only the spring of 2012)---------

wst11 <- d3 %>% 
  filter(DateTimeUTC > "2011-06-01 00:00:00", DateTimeUTC < '2012-05-31 23:59:59') # keep in mind that this is all in UTC still

fp_wst11 <- fishpaths(wst11, wst11$TagID, wst11$Station)

fp_wst11 <- fp_wst11 %>% 
  arrange(arrival) %>% 
  mutate(residence = departure - arrival)
  
units(fp_wst11$residence) <- "days"

fp_wst11 %>% 
  filter(residence <= 7) # filter out encounters equal to or longer than 7 consecutive days

firstlast_wst11 <- fp_wst11 %>% 
  group_by(TagID) %>% 
  summarise(firstarrival = min(arrival), lastdeparture = max(departure)) %>% 
  mutate(totalres = lastdeparture - firstarrival) 

units(firstlast_wst11$totalres) <- "days"
firstlast_wst11$tagging_group <- "wst11"

meanres_wst11 <- firstlast_wst11 %>% 
  summarise(meanres = mean(totalres), sdres = sd(totalres), n = len(TagID)) # works; move to new script to complete calculations

meanres_wst11


```
```{r}
# White sturgeon residence in 2012------------------

wst12 <- d3 %>% 
  filter(DateTimeUTC > "2012-06-01 00:00:00", DateTimeUTC < '2013-05-31 23:59:59') # keep in mind that this is all in UTC still

fp_wst12 <- fishpaths(wst12, wst12$TagID, wst12$Station)

fp_wst12 <- fp_wst12 %>% 
  arrange(arrival) %>% 
  mutate(residence = departure - arrival)
  
units(fp_wst12$residence) <- "days"

fp_wst12 %>% 
  filter(residence <= 7) # filter out encounters equal to or longer than 7 consecutive days

firstlast_wst12 <- fp_wst12 %>% 
  group_by(TagID) %>% 
  summarise(firstarrival = min(arrival), lastdeparture = max(departure)) %>% 
  mutate(totalres = lastdeparture - firstarrival) 

units(firstlast_wst12$totalres) <- "days"
firstlast_wst12$tagging_group <- "wst12"

meanres_wst12 <- firstlast_wst12 %>% 
  summarise(meanres = mean(totalres), sdres = sd(totalres), n = len(TagID)) # works; move to new script to complete calculations

meanres_wst12


```
```{r}
# White sturgeon residence in 2013:-----------------
wst13 <- d3 %>% 
  filter(DateTimeUTC > "2013-06-01 00:00:00", DateTimeUTC < '2014-05-31 23:59:59')

wst13 <- filter(wst13,  !(TagID %in% tags$wst_2014.TagID)) # Only run this line for the plot; it filters out the fish tagged that spring

fp_wst13 <- fishpaths(wst13, wst13$TagID, wst13$Station)

fp_wst13 <- fp_wst13 %>% 
  arrange(arrival) %>% 
  mutate(residence = departure - arrival)
  
units(fp_wst13$residence) <- "days"

fp_wst13 %>% 
  filter(residence <= 7) # filter out encounters equal to or longer than 7 consecutive days

firstlast_wst13 <- fp_wst13 %>% 
  group_by(TagID) %>% 
  summarise(firstarrival = min(arrival), lastdeparture = max(departure)) %>% 
  mutate(totalres = lastdeparture - firstarrival) 

units(firstlast_wst13$totalres) <- "days"
firstlast_wst13$tagging_group <- "wst13"

meanres_wst13 <- firstlast_wst13 %>% 
  summarise(meanres = mean(totalres), sdres = sd(totalres), n = len(TagID)) # works; move to new script to complete calculations

meanres_wst13

```
```{r}
# White sturgeon residence in 2014:-----------------
wst14 <- d3 %>% 
  filter(DateTimeUTC > "2014-06-01 00:00:00", DateTimeUTC < '2015-05-31 23:59:59') # keep in mind that this is all in UTC still

fp_wst14 <- fishpaths(wst14, wst14$TagID, wst14$Station)

fp_wst14 <- fp_wst14 %>% 
  arrange(arrival) %>% 
  mutate(residence = departure - arrival)
  
units(fp_wst14$residence) <- "days"

fp_wst14 %>% 
  filter(residence <= 7) # filter out encounters equal to or longer than 7 consecutive days

firstlast_wst14 <- fp_wst14 %>% 
  group_by(TagID) %>% 
  summarise(firstarrival = min(arrival), lastdeparture = max(departure)) %>% 
  mutate(totalres = lastdeparture - firstarrival) 

units(firstlast_wst14$totalres) <- "days"
firstlast_wst14$tagging_group <- "wst14"

meanres_wst14 <- firstlast_wst14 %>% 
  summarise(meanres = mean(totalres), sdres = sd(totalres), n = len(TagID)) # works; move to new script to complete calculations

meanres_wst14

```

```{r}
# White sturgeon residence in 2015-2016:-----------------
wst15 <- d3 %>% 
  filter(DateTimeUTC > "2015-06-01 00:00:00", DateTimeUTC < '2016-05-31 23:59:59') # keep in mind that this is all in UTC still

fp_wst15 <- fishpaths(wst15, wst15$TagID, wst15$Station)

fp_wst15 <- fp_wst15 %>% 
  arrange(arrival) %>% 
  mutate(residence = departure - arrival)
  
units(fp_wst15$residence) <- "days"

fp_wst15 %>% 
  filter(residence <= 7) # filter out encounters equal to or longer than 7 consecutive days

firstlast_wst15 <- fp_wst15 %>% 
  group_by(TagID) %>% 
  summarise(firstarrival = min(arrival), lastdeparture = max(departure)) %>% 
  mutate(totalres = lastdeparture - firstarrival) 

units(firstlast_wst15$totalres) <- "days"
firstlast_wst15$tagging_group <- "wst15"

meanres_wst15 <- firstlast_wst15 %>% 
  summarise(meanres = mean(totalres), sdres = sd(totalres), n = len(TagID)) # works; move to new script to complete calculations

meanres_wst15
```

```{r}
# wst residence
allwst <- rbind(firstlast_wst11, firstlast_wst12, firstlast_wst13, firstlast_wst14, firstlast_wst15)
allwst$Sp = "wst"

ggplot(allwst, aes(x = Sp, y = totalres)) + geom_boxplot(alpha = 0.5) + geom_jitter(aes(color = tagging_group))
```



```{r}

# How to filter for shed tags:
# we want to exclude any tags that pinged continuously for more than 1 week, which means taht we have to filter them before the fish paths take place.
# given d, spit out the TagID of any fish that sat in once place had consecutive detections.  Could do this with cumsum?

# if a tag pings every 30 seconds for a week, what would the cumsum be on the difftime?
# another way to ask that: how many 30-second intervals are there in a week?  
intervals = 604800/30 # about 20,000 detections.  So need a way to filter out encounters that have >20,000 continuous detections.  First let's find any:

d2 <- filter(d, TagID %in% tags$fca_2014.TagID)
d2 <- filter(d2, DateTimeUTC >= max(d2$DateTagged))
d2 %>% 
  group_by(TagID, Station) %>% 
  summarise(n = n()) %>% 
  arrange(-n) %>% 
  top_n(10)

# so a fish died at Cache_creek and at Wallace Weir, but let's make sure that was their final detected location

shed_cache <- filter(d, TagID == 13729)
shed_fp <- fishpaths(shed_cache, shed_cache$TagID, shed_cache$Station) # true for this guy, although he also has a weird detection before he was tagged - need to make that a default filter somehow.

shed_ww <- filter(d, TagID == 20164)
shed_wwfp <- fishpaths(shed_ww, shed_ww$TagID, shed_ww$Station) # true for this guy as well.

# Ok - so proof of concept, now we need to filter out just those detections.

d2_filt <- d2 %>% 
  group_by(TagID, Station) %>% 
  mutate(count = n()) %>% 
  filter(count <= 17000) %>% 
  ungroup() %>% 
  arrange(DateTimeUTC)

13729 %in% d2_filt$TagID # TRUE, should still be in there, just not at Cache_creek
shed_cache_chk <- filter(d2_filt, TagID == 13729)
shed_fp_chk <- fishpaths(shed_cache_chk, shed_cache_chk$TagID, shed_cache_chk$Station) # this kind of works, except that we lost that location info - the fish DID make it to Cache Creek, and we want that reflected.


ggplot(d2, aes(x = DateTimeUTC, y = Station)) + geom_point(aes(color = factor(TagID)))
#jesus, how long was that tag there?

d2 %>% 
  filter(TagID == 13729) %>% 
  arrange(DateTimeUTC) %>% 
  group_by(Station) %>% 
  summarise(timespent = max(DateTimeUTC) - min(DateTimeUTC))

# same, but on fishpaths df:
shed_fp <- shed_fp %>% 
  arrange(arrival) %>% 
  mutate(residence = departure - arrival)
  
units(shed_fp$residence) <- "days"
shed_fp

shed_fp %>% 
  filter(residence <= 7) # filter out encounters equal to or longer than 7 consecutive days # Okay - that should work.
```

```{r}
# Chinook residence in 2012-2013:
d2 <- filter(d, TagID %in% tags$fca_2012.TagID)
d2 <- d2 %>% 
  group_by(TagID) %>% 
  filter(DateTimeUTC >= DateTagged) # make sure we don't get any detections for individual fish before they were tagged

chn12 <- d2 %>% 
  filter(DateTimeUTC > "2012-06-01 00:00:00", DateTimeUTC < '2013-05-31 23:59:59') # keep in mind that this is all in UTC still

fp_chn12 <- fishpaths(chn12, chn12$TagID, chn12$Station)

fp_chn12 <- fp_chn12 %>% 
  arrange(arrival) %>% 
  mutate(residence = departure - arrival)
  
units(fp_chn12$residence) <- "days"

fp_chn12 %>% 
  filter(residence <= 7) # filter out encounters equal to or longer than 7 consecutive days

firstlast_chn12 <- fp_chn12 %>% 
  group_by(TagID) %>% 
  summarise(firstarrival = min(arrival), lastdeparture = max(departure)) %>% 
  mutate(totalres = lastdeparture - firstarrival) 

units(firstlast_chn12$totalres) <- "days"
firstlast_chn12$tagging_group <- "chn12"

meanres_chn12 <- firstlast_chn12 %>% 
  summarise(meanres = mean(totalres), sdres = sd(totalres), n = len(TagID)) 

meanres_chn12

# Chinook residence in 2013-2014
d2 <- filter(d, TagID %in% tags$fca_2013.TagID)

d2 <- d2 %>% 
  group_by(TagID) %>% 
  filter(DateTimeUTC >= DateTagged) %>% 
  ungroup() # make sure we don't get any detections for individual fish before they were tagged; filters out 702 detections

chn13 <- d2 %>% 
  filter(DateTimeUTC > "2013-06-01 00:00:00", DateTimeUTC < '2014-05-31 23:59:59') %>% 
  arrange(DateTimeUTC)# keep in mind that this is all in UTC still

fp_chn13 <- fishpaths(chn13, chn13$TagID, chn13$Station)

fp_chn13 <- fp_chn13 %>% 
  arrange(arrival) %>% 
  mutate(residence = departure - arrival)
  
units(fp_chn13$residence) <- "days"

fp_chn13 <- fp_chn13 %>% 
  filter(residence <= 7) # filter out encounters equal to or longer than 7 consecutive days

firstlast_chn13 <- fp_chn13 %>% 
  group_by(TagID) %>% 
  summarise(firstarrival = min(arrival), lastdeparture = max(departure)) %>% 
  mutate(totalres = lastdeparture - firstarrival) 

units(firstlast_chn13$totalres) <- "days"
firstlast_chn13$tagging_group <- "chn13"

meanres_chn13 <- firstlast_chn13 %>% 
  summarise(meanres = mean(totalres), sdres = sd(totalres), n = len(TagID)) 

meanres_chn13

# Chinook residence in 2014-2015-----------------

d2 <- filter(d, TagID %in% tags$fca_2014.TagID)

d2 <- d2 %>% 
  group_by(TagID) %>% 
  filter(DateTimeUTC >= DateTagged) %>% 
  ungroup() # make sure we don't get any detections for individual fish before they were tagged

chn14 <- d2 %>% 
  filter(DateTimeUTC > "2014-06-01 00:00:00", DateTimeUTC < '2015-05-31 23:59:59') %>% 
  arrange(DateTimeUTC)# keep in mind that this is all in UTC still

fp_chn14 <- fishpaths(chn14, chn14$TagID, chn14$Station)

fp_chn14 <- fp_chn14 %>% 
  arrange(arrival) %>% 
  mutate(residence = departure - arrival)
  
units(fp_chn14$residence) <- "days"

fp_chn14 <- fp_chn14 %>% 
  filter(residence <= 7) # filter out encounters equal to or longer than 7 consecutive days


firstlast_chn14 <- fp_chn14 %>% 
  group_by(TagID) %>% 
  summarise(firstarrival = min(arrival), lastdeparture = max(departure)) %>% 
  mutate(totalres = lastdeparture - firstarrival) 

units(firstlast_chn14$totalres) <- "days"
firstlast_chn14$tagging_group <- "chn14"

meanres_chn14 <- firstlast_chn14 %>% 
  summarise(meanres = mean(totalres), sdres = sd(totalres), n = len(TagID)) 

meanres_chn14

ggplot(fp_chn14, aes(x = departure, y = Station)) + geom_point(aes(color = factor(TagID))) # so the problem is that separately, none of the encounters are more than 7 days, but cumulatively, there's still definnitely a couple of potential shed tags.  Will need to come up with a better, more subtle filter for this down the road.  # GIT ISSUE


# Chinook residence in 2015-2016 ------------------------------

d2 <- filter(d, TagID %in% tags$fca_2015.TagID)

d2 <- d2 %>% 
  group_by(TagID) %>% 
  filter(DateTimeUTC >= DateTagged) %>% 
  ungroup() # make sure we don't get any detections for individual fish before they were tagged

chn15 <- d2 %>% 
  filter(DateTimeUTC > "2015-06-01 00:00:00", DateTimeUTC < '2016-05-31 23:59:59') %>% 
  arrange(DateTimeUTC)# keep in mind that this is all in UTC still

fp_chn15 <- fishpaths(chn15, chn15$TagID, chn15$Station)

fp_chn15 <- fp_chn15 %>% 
  arrange(arrival) %>% 
  mutate(residence = departure - arrival)
  
units(fp_chn15$residence) <- "days"

fp_chn15 <- fp_chn15 %>% 
  filter(residence <= 7) # filter out encounters equal to or longer than 7 consecutive days

firstlast_chn15 <- fp_chn15 %>% 
  group_by(TagID) %>% 
  summarise(firstarrival = min(arrival), lastdeparture = max(departure)) %>% 
  mutate(totalres = lastdeparture - firstarrival) 

units(firstlast_chn15$totalres) <- "days"
firstlast_chn15$tagging_group <- "chn15"

meanres_chn15 <- firstlast_chn15 %>% 
  summarise(meanres = mean(totalres), sdres = sd(totalres), n = len(TagID)) 

meanres_chn15
ggplot(fp_chn15, aes(x = departure, y = Station)) + geom_point(aes(color = factor(TagID)))

```

So now, after doing all those clunky brute-force calculations, time to put it all together:

```{r}
chn <- rbind(firstlast_chn12, firstlast_chn13, firstlast_chn14, firstlast_chn15)
chn$Sp <- "chn"

residency <- rbind(allwst, chn)

res_completeyears <- filter(residency, tagging_group != "wst11", tagging_group != "chn12")
g <- ggplot(res_completeyears, aes(x = Sp, y = totalres)) + geom_boxplot(alpha = 0.6) + geom_jitter(aes(color = Sp), size = 2, alpha = 0.7)

g + labs(title = "Residency in Yolo Bypass by Species", x = "Species", y = "Residence in Days")

```
This data has been lightly filtered for shed tags, and for consistency in length of year and tagging location, it does not include the spring of 2012, the values for the white sturgeon tagged in Spring of 2014 within the "wst13 group", or the 12 salmon tagged in fall of 2012.  A complete table of average residencies by species and year (including those groups ommited above) is below.

```{r}

residency %>% 
  group_by(tagging_group) %>% 
  summarise(`Mean Residency in Days` = mean(totalres), `Standard Deviation` = sd(totalres), `Number of Fish Tagged` = n())

```
